\chapter{Changelog}

\section{GnuCOBOL 3.0-dev}

\paragraph{General}
\begin{itemize}
\item \textbf{C API (configuration)}: added cob\_set\_runtime\_option and cob\_get\_runtime\_option for specifying the file handle which printer output or trace output should go to.
\item \textbf{cobc command-line options}: added -fdump to dump data of all modules on abend of program.
\item \textbf{cobc command-line options}: disabled -fif-cutoff pending removal.
\item \textbf{cobc command-line options}: -debug no longer implies -ftrace (\bug{449}).
\item \textbf{cobc command-line options}: added -O0 for disabling optimisations (\fr{255}).
\item \textbf{cobc command-line options}: added -Wno-dialect to suppress dialect-specific warnings.
\item \textbf{cobc command-line options}: added -Wother to enable miscellaneous warnings.
\item \textbf{cobcrun command-line options}: renamed -runtime-conf to -runtime-config.
\item \textbf{Copyfiles}: fixed segfault when GnuCOBOL tried and failed to open a copyfile (\bug{458}).
\item \textbf{Copyfiles}: GnuCOBOL now checks for recursive copyfiles (\bug{467}).
\item \textbf{Debugging}: Added debug logging for compiler developers (--enable-debug-log ./configure option).
\item \textbf{DJGGP support}: added minimal support for DJGGP.
\item \textbf{Expressions}: fixed \bug{431}, where decimal constants were not initialised in INITIAL programs.
\item \textbf{IBM i support}: fixed segfault involving filepaths.
\item \textbf{Listings}: fixed the substitution of tokens in quotes in copyfiles (\bug{494}).
\item \textbf{Modules}: enabled running of modules generated with older versions of GnuCOBOL (\fr{239}).
\item \textbf{Nested programs}: fixed \bug{435}, where identifiers in the containing program where not propogated to the contained program, causing a segfault.
\item \textbf{Recursion}: fixed \bug{222}, where the return value was lost when returning from a recursive call.
\item \textbf{Reserved words}: allow users to make default reserved words aliases for other words.
\item \textbf{Signals}: added error message for SIGFPE (see \bug{434}).
\item \textbf{Solaris support}: fixed errors and warnings when compiling in Solaris 10.
\item \textbf{Tracing}: added new options for controlling tracing (see \fr{242}).
\end{itemize}

\paragraph{Configuration options}
\begin{itemize}
\item \textbf{IBM dialect}: reserved words updated to Enterprise COBOL V6.2.
\item \textbf{incorrect-conf-sec-order option}: changed to ``ok'' from ``error'' in mf and default dialects.
\item \textbf{New compiler configuration options}: binary-comp-1, display-special-fig-consts, free-redefines-position (\fr{211}), line-col-zero-default, move-figurative-space-to-numeric, move-non-numeric-lit-to-numeric-is-zero, missing-statement, perform-without-varying-by, record-delimiter, record-delim-with-fixed-recs, screen-section-rules, sequential-delimiters
\item \textbf{New runtime configuration options}: col\_just\_lrc (for enabling\slash{}disabling the LEFT\slash{}RIGHT\slash{}CENTER phrases of the report COLUMN clause), printer and display\_print\_pipe (for specifying what command should be executed before DISPLAY UPON PRINTER; similar to Micro Focus' COBPRINTER option); display\_print\_file (name of file which DISPLAY UPON PRINTER will append its output to), trace\_format, dump\_file and dump\_width.
\item \textbf{Registers}: disabling a register also removes the reserved word with the same name, if it exists (\fr{278}).
\end{itemize}

\paragraph{Compiler directives}
\begin{itemize}
\item \textbf{New directives}: ADDRSV, ADDSYN, MAKESYN, OVERRIDE and REMOVE (\fr{210}); COMP1.
\end{itemize}

\paragraph{Identification division}
\begin{itemize}
\item \textbf{OPTIONS paragraph}: added support for ARITHMETIC clause.
\end{itemize}

\paragraph{Environment division}
\begin{itemize}
\item \textbf{ALTERNATE RECORD KEY clause}: implemented split keys (SOURCE IS; \fr{23}) and sparse keys (SUPPRESS WHEN; \fr{281}).
\item \textbf{ASSIGN clause}: re-enabled use of linkage section or BASED items (\bug{421}).
\item \textbf{CLASS phrase}: added recognition of ALPHANUMERIC\slash{}NATIONAL and alphabet phrases.
\item \textbf{CURRENCY phrase}: changed to emit error when CURRENCY SIGN other than ``\$'' is entered.
\item \textbf{CURRENCY phrase}: improved error messages for invalid currency signs.
\item \textbf{EXTERN clause}: improved error messages (\bug{446}).
\item \textbf{FILE STATUS clause}: added detection of VSAM secondary status identifier (\fr{51}).
\item \textbf{OCCURS clause}: correctly implemented nested OCCURS DEPENDING tables (ODOSLIDE).
\item \textbf{PASSWORD clause}: added detection of.
\item \textbf{RECORD DELIMITER phrase}:  added support for BINARY-SEQUENTIAL and LINE-SEQUENTIAL phrases.
\item \textbf{RECORD DELIMITER phrase}: improved syntax checks (see \bug{442}).
\item \textbf{RECORD KEY clause}: implemented sparse keys (SOURCE IS; \fr{23}).
\end{itemize}

\paragraph{Data division}
\begin{itemize}
\item \textbf{ANY LENGTH clause}: fixed \bug{487}, where literals moved to ANY LENGTH items were incorrectly truncated.
\item \textbf{Screen section}: made some COBOL words context-sensitive to screen section.
\item \textbf{Screen description}: the rules on which clauses must be specified when now depend on the dialect (see \bug{382}).
\item \textbf{SYNCHRONIZED clause}: deactivated RIGHT phrase, pending correct implementation (previously SYNCHRONIZED RIGHT was the same as SYNCHRONIZED LEFT).
\item \textbf{USAGE clause}: added option to make COMP-1 mean a 16-bit signed integer (\fr{272}).
\end{itemize}

\paragraph{Procedure division}
\begin{itemize}
\item \textbf{ACCEPT statement (screen)}: fixed ACCEPT WITH UPDATE and not working (\bug{423}).
\item \textbf{ACCEPT statement (screen)}: fixed \bug{426}, where backspacing at the start of a field moved the cursor to the second-to-last character of the preceding field.
\item \textbf{ACCEPT statement (screen)}: added detection of CONTROL KEY clause.
\item \textbf{ACCEPT statement (screen)}: allow numeric-edited fields to contain spaces (see \bug{491}).
\item \textbf{ACCEPT statement (screen)}: fixed error when filling in a one character field with insert mode on (\bug{498}).
\item \textbf{ACCEPT statement (temporal)}: fixed \bug{469}, where ACCEPT FROM DAY was off by $-1$.
\item \textbf{CALL statement}: fixed ON EXCEPTION not working properly with -fstatic (\bug{462}).
\item \textbf{DISPLAY statement (printer)}: allow for redirecting DISPLAY UPON PRINTER(-1) to files based on the runtime configuration.
\item \textbf{DISPLAY statement (screen)}: fixed DISPLAY LOW-VALUE not setting position of cursor for next DISPLAY statement (\bug{423}).
\item \textbf{DISPLAY statement (screen)}: added configuration option to disable Micro Focus' special behaviour with some figurative constants (see \bug{423}).
\item \textbf{DISPLAY statement (screen)}: fixed \bug{428}, where DISPLAY ALL ``x'' WITH SIZE only displayed ``x'' once.
\item \textbf{Exceptions}: the exception status is now only reset by SET LAST EXCEPTION TO OFF.
\item \textbf{Exceptions}: fixed EC-SIZE-OVERFLOW being raised when EC-SIZE-ZERO-DIVIDE is active (\bug{223}).
\item \textbf{File I-O}: fixed \bug{457}, where file status 30 was set instead of the correct permanent error status (34, 35 or 37).
\item \textbf{File I-O}: added support for VBISAM 2.1.1.
\item \textbf{Floating-point arithmetic}: fixed incorrect SIZE ERROR exceptions (see \bug{470}).
\item \textbf{Floating-point arithmetic}: fixed \bug{478}, where calculations with the SIZE ERROR phrase have different results to calculations without it.
\item \textbf{Floating-point arithmetic}: now activate the SIZE ERROR handler when a floating-point variable is set to a non-finite value (see \bug{122}).
\item \textbf{INQUIRE statement}: added detection of.
\item \textbf{LENGTH OF phrase}: fixed error message when used on group fields (\bug{175}).
\item \textbf{PERFORM statement}: added check for non-zero item in BY phrase (\fr{268}).
\item \textbf{PERFORM statement}: allow the BY phrase to be omitted, which is the same as specifying BY 1 (\fr{158}).
\item \textbf{MODIFY statement}: added detection of.
\item \textbf{MOVE statement}: added options to interpret moving non-numeric values to numeric items as moving zero to numeric items.
\item \textbf{Reference modification}: improved out-of-bounds error message (\bug{445}).
\item \textbf{OPEN statement}: fail with file status 39 when the first record of an indexed file is larger than specified in the FD.
\item \textbf{READ statement (indexed)}: fail with file status 43 when the read record is larger than specified in the FD.
\item \textbf{Reference modification}: added warnings for invalid reference modifications involving variables.
\item \textbf{Report writer}: implemented, including features from COBOL 2002 and the IBM Report Writer.
\item \textbf{SET statement}: fixed \bug{225}, where an invalid SET statement caused an error saying ``invalid MOVE statement''.
\item \textbf{STOP statement}: fixed the NORMAL phrase causing a compilation error (\bug{433}).
\item \textbf{VALIDATE statement}: added detection of.
\end{itemize}

\paragraph{Intrinsic functions}
\begin{itemize}
\item \textbf{MOD function}: fixed EC-SIZE-ZERO-DIVIDE being raised instead of EC-ARGUMENT-FUNCTION when zero was provided as an argument.
\item \textbf{REM function}: fixed EC-SIZE-ZERO-DIVIDE being raised instead of EC-ARGUMENT-FUNCTION when zero was provided as an argument.
\item \textbf{WHEN-COMPILED function}: fixed timezone, which was missing its sign and contained nonsense when negative (\bug{436}).
\end{itemize}

\paragraph{Built-in subprograms}
\begin{itemize}
\item \textbf{C\$GETPID}: fixed wrong process ID being returned after forking (\bug{451}).
\item \textbf{C\$SLEEP}: if the requested time is too large, sleep for the maximum possible time instead of not at all.
\item \textbf{CBL\_READ\_KBD\_CHAR}: this now works (\bug{500}).
\item \textbf{SYSTEM}: added workaround of buggy system() implementation on Windows, which removes leading and trailing quotes.
\end{itemize}

\section{GnuCOBOL 2.2}

This list tracks changes made from 23 November 2013. This excludes many changes made in 2009--2013 which would be pertinent to those upgrading from a 2009 build of OpenCOBOL 1.1 found in many package repositories.

\paragraph{General}
\begin{itemize}
\item \textbf{64-bit numbers}: fixed bugs in handling of 64-bit numbers (e.g. \bug{229}).
\item \textbf{ACUCOBOL windows}: added detection of ACUCOBOL's window\slash{}message box GUI syntax.
\item \textbf{C API (data)}: added several functions for getting and setting cob\_field items.
\item \textbf{C API (files)}: added cob\_file\_external\_addr, cob\_file\_malloc and cob\_file\_free.
\item \textbf{C API (screen)}: added several functions from Micro Focus' C to COBOL API: cob\_display\_text, cob\_sys\_get\_char, cob\_get\_char, cob\_get\_text, cob\_display\_formatted\_text, cob\_sys\_get\_csr\_pos, cob\_sys\_set\_csr\_pos, cobmove, cobaddstrc, cobprintf and cobgetch (feature requests \frnoname{148} and~\frnoname{187}).
\item \textbf{C API (signals)}: added cob\_raise to send signal to signal handlers.
\item \textbf{C compiler support}: fixed errors in compilers without designated initializers.
\item \textbf{cobc command-line options}: added -O3 to enable more optimisations.
\item \textbf{cobc command-line options}: added -Wfatal-error to make the compiler abort on the first error.
\item \textbf{cobc command-line options}: added -Wpossible-overlap to warn items that \emph{may} overlap (-Woverlap only warns if items definitely overlap).
\item \textbf{cobc command-line options}: added -fmax-errors to set number of errors at which the compiler aborts.
\item \textbf{cobc command-line options}: added -fwinmain to output WinMain instead of main (\fr{194}).
\item \textbf{cobc command-line options}: added -t and -T for complete listing support (-t for 80-characters wide listings and -T for 132-characters wide) which includes cross-references (thanks to Dave Pitts).
\item \textbf{cobc command-line options}: added -vvv (like -vv but passes verbose option to the linker as well) and -\#\#\# (like -v but commands are not executed).
\item \textbf{cobc command-line options}: allow -, i.e. stdin, as a source file.
\item \textbf{COBOL-85 NIST testsuite}: tests now refer to \$COBC, \$COBCRUN and \$COBCRUN\_DIRECT environment variables instead of directly calling cobc and cobcrun, allowing the testsuite to run in conjunction with tools like valgrind.
\item \textbf{COBOL-85 NIST testsuite}: tests for obsolete feature flagging are now executed, if possible.
\item \textbf{Comments}: added ACUCOBOL comments: \$ as synonym for * in indicator area and $\vert$ as synonym for floating comment indicator *>.
\item \textbf{Communication facility}: added detection of communication facility syntax.
\item \textbf{configure}: added useful error message when help2man, bison and flex are missing when they are needed.
\item \textbf{curses}: fixed compilation errors when configured without curses (\bug{90}).
\item \textbf{Error messages}: error messages are now lowercase, in line with the GNU Coding Standards (\bug{198}).
\item \textbf{Error messages}: segfaults in the compiler now cause an error message to be displayed.
\item \textbf{Error messages}: replaced instances of ``ODO'' by the clearer ``OCCURS DEPENDING ON''.
\item \textbf{Expressions}: resolve constant expressions and optimise constant decimals at compile time.
\item \textbf{Expressions}: added support for IBM OS/VS COBOL's arithmetic.
\item \textbf{Expressions}: improved error messages for malformed expressions.
\item \textbf{Indicators}: invalid indicators no longer cause compilation to immediately terminate (\fr{126}).
\item \textbf{Information}: output compiler version used to build GnuCOBOL and any mathematical libraries used (\fr{169}).
\item \textbf{Information}: output what a reserved word is an alias for in the --list-reserved output (\fr{214}).
\item \textbf{Manpage}: added manpage generation and installation.
\item \textbf{Nested programs}: Nested programs no longer need to have END PROGRAM.
\item \textbf{National literals}: added basic support for national literals.
\item \textbf{Numeric literals}: added ACUCOBOL numeric literals: B\#\ldots\, for binary, O\#\ldots\, for octal, and X\#\ldots\ and H\#\ldots\, for hexadecimal.
\item \textbf{Literals}: fixed heap corruptions caused by uncommon literals (\bug{195}).
\item \textbf{Literals}: allow concatenation of literal and Boolean literals.
\item \textbf{Memory management}: all memory belonging to the parsers and lexers is freed upon a compiler abend.
\item \textbf{Memory management}: fixed memory leaks due to recursive CALLs.
\item \textbf{Microsoft Visual C++}: output when compiling with cl.exe is now filtered and temporary files are deleted.
\item \textbf{MinGW}: fixed use of wrong directory separator.
\item \textbf{Signals}: removed error message on SIGPIPE.
\item \textbf{Signals}: added error message for SIGBUS.
\item \textbf{Translations}: updated, with new support for German and Italian.
\item \textbf{User-defined functions}: function definitions must now end with END FUNCTION.
\item \textbf{User-defined functions}: function definitions may no longer be nested in programs (\bug{255}).
\item \textbf{Windows support}: allow linking with asm files.
\item \textbf{Windows support}: added support for DISAM in the batch file which creates distributables.
\item \textbf{Windows support}: fixed environment-setting batch files not working with Microsoft Visual Studio 2017.
\item \textbf{Windows support}: fixed 64-bit environment-setting batch files not checking the correct directories for binaries and libraries.
\end{itemize}

\paragraph{Configuration options}
\begin{itemize}
\item \textbf{Deleted compiler configuration options}: eject-statement, cobol85-reserved.
\item \textbf{New compiler configurations}: all dialects have been split into standard and strict dialects, with strict dialects maintaining source compatibility with the dialect's compiler(s).
\item \textbf{New compiler configurations}: acu for ACUCOBOL, cobol2014 for COBOL 2014, rm for RM-COBOL, xopen for X\slash{}Open.
\item \textbf{New compiler configuration options}: accept-display-extensions, accept-update, accept-auto, acu-literals, arithmetic-osvs, call-overflow, console-is-crt, constant-01, constant-78, constant-folding, define-constant-directive, hexadecimal-boolean, hexadecimal-national-literals, incorrect-conf-sec-order, intrinsic-function, listing-statements, literal-length, move-figurative-constant-to-numeric, move-figurative-quote-to-numeric, move-ibm, national-literals, no-echo-means-secure, not-exception-before-exception, numeric-boolean, numeric-literal-length, numeric-value-for-edited-item, pic-length, program-name-redefinition, program-prototypes, reference-out-of-declaratives (\fr{179}), register, renames-uncommon-levels, reserved, reserved-words, stop-identifier, system-name, title-statement, use-for-debugging, word-length (\fr{43}). % TO-DO: expand
\item \textbf{Registers}: compiler configurations can now specify all the registers to generate.
\item \textbf{Registers}: added registers not yet implemented by GnuCOBOL as reserved words.
\item \textbf{Renamed compiler configuration options}: debugging-line to debugging-mode, relaxed-syntax-check to relax-syntax-checks.
\item \textbf{Reserved words}: compiler configurations can now specify all the reserved words and context-sensitive words permitted.
\item \textbf{Reserved words}: compiler configurations can now specify whether a reserved word is an alias for another reserved word. % TO-DO: Improve
\item \textbf{Runtime configuration}: added ability to configure some libcob features at runtime.
\item \textbf{Support options}: options which specify if a feature is supported can now take a ``+'' before their argument to indicate it takes effect only if the current level of support is less strict than ``ok''.
\end{itemize}

\paragraph{Compiler directives}
\begin{itemize}
\item \textbf{\$ indicator character}: added \$ as an indicator for compiler directive lines.
\item \textbf{\directiveindicator{}IF directive}: fixed \bug{263}, where nested \directiveindicator{}IF directives were not handled correctly.
\item \textbf{New constants}: GCCOMP, GNUCOBOL.
\item \textbf{New directives}: \directiveindicator{}CALL-CONVENTION, \directiveindicator{}LISTING, \directiveindicator{}PAGE. % TO-DO: Expand
\item \textbf{New directives (detection only)}: *CBL, *CONTROL, TITLE.
\item \textbf{New \directiveindicator{}SET phrase}: SOURCEFORMAT.
\end{itemize}

\paragraph{Identification division}
\begin{itemize}
\item \textbf{Comment paragraphs}: fixed invalid parsing of quote characters inside comment paragraphs (\bug{297}).
\item \textbf{FUNCTION-ID}: added checks for redefinition of function-names.
\item \textbf{INITIAL phrase}: fixed premature deallocation of INITIAL programs (\bug{52}).
\item \textbf{OPTIONS paragraph}: added with implementation of DEFAULT ROUNDED MODE and ENTRY-CONVENTION phrases and recognition of INTERMEDIATE ROUNDING phrase.
\item \textbf{PROGRAM-ID}: added checks for redefinition of program-names.
\item \textbf{PROGRAM-ID phrases}: permit INITIAL or RECURSIVE before COMMON (\bug{244}).
\item \textbf{Program\slash{}function-names}: warn if program\slash{}function-names contain spaces.
\end{itemize}

\paragraph{Environment division}
\begin{itemize}
\item \textbf{ASSIGN clause}: missing ASSIGN clauses are now detected at compile-time.
\item \textbf{ASSIGN clause}: added PRINTER and PRINTER-1 device-names for writing to a printer.
\item \textbf{ASSIGN clause}: added CARD-PUNCH, CARD-READER, CASSETTE, INPUT, INPUT-OUTPUT, MAGNETIC-TAPE and OUTPUT device-names for line sequential devices.
\item \textbf{ASSIGN clause}: temporarily prohibit BASED and linkage items in ASSIGN USING due to \bug{421}.
\item \textbf{CALL-CONVENTION phrase}: statically calling functions with CALL-CONVENTION 74 no longer causes linker errors (\bug{316}).
\item \textbf{CURRENCY phrase}: fixed \bug{182}, where a preceding SWITCH phrase caused an incorrect duplicate CURRENCY clause error.
\item \textbf{File-control entry}: fixed \bug{71}, where referring to a global constant caused an internal error.
\item \textbf{File-control entry}: fixed \bug{331}, where using an identifier in a file record qualified with the file's name caused an error.
\item \textbf{FUNCTION phrase}: added checks for redefinition of function-(prototype-)names.
\item \textbf{FUNCTION phrase}: compiler will no longer stop when it encounters a syntax error.
\item \textbf{LOCK MODE clause}: fixed combination of LOCK MODE IS AUTOMATIC/MANUAL with LOCK ON MULTIPLE.
\item \textbf{PROGRAM phrase}: added support for program-prototype-names.
\item \textbf{SIGN clause}: improved syntax checks.
\item \textbf{SWITCH phrase}: added check for duplicate on\slash{}off clauses (\bug{136}).
\item \textbf{SWITCH phrase}: added new switch names: SWITCH-16 through to SWITCH-36 (\fr{65}), ``SWITCH 1'' to ``SWITCH 26'' (and their aliases  ``SWITCH A'' to ``SWITCH Z''), UPSI-0 to UPSI-8 (equivalent to ``SWITCH 0'' to ``SWITCH 8'') and USW-0 to USW-31 (equivalent to ``SWITCH 0'' to ``SWITCH 31'').
\end{itemize}

\paragraph{Data division}
\begin{itemize}
\item \textbf{78-level items}: strengthened syntax checks.
\item \textbf{88-level items}: strengthened syntax checks.
\item \textbf{ANY NUMERIC clause}: ANY NUMERIC items must now have PIC 9.
\item \textbf{ANY LENGTH clause}: ANY LENGTH items may no longer be BY VALUE parameters (see \bug{219}).
\item \textbf{ANY LENGTH clause}: ANY LENGTH items must now have PIC X or PIC N.
\item \textbf{BLANK clause}: fixed \bug{143}, where BLANK LINE\slash{}SCREEN did not colour line\slash{}screen.
\item \textbf{BLANK WHEN ZERO clause}: added checks that BLANK WHEN ZERO is not specified with PICTURE clauses containing S.
\item \textbf{Constant items}: expressions in VALUE clauses now permitted.
\item \textbf{Data description}: added a maximum record length.
\item \textbf{Data description}: increased maximum size of non-indexed file record to 64 MiB (maximum size of an indexed file record is 65535 bytes).
\item \textbf{ERASE clause}: fixed \bug{186}, where ERASE EOL and ERASE EOS could be specified simultaneously.
\item \textbf{FULL clause}: added warning for useless FULL clauses on numeric items (\fr{209}).
\item \textbf{HIGHLIGHT and LOWLIGHT clauses}: added checks that HIGHLIGHT and LOWLIGHT are not specified simultaneously.
\item \textbf{Local-storage section}: fixed \bug{78}, where local-storage items where initialised after file section items.
\item \textbf{LOWLIGHT clause}: implemented.
\item \textbf{OCCURS clause}: fixed internal compiler when used with SYNC (\bug{155}).
\item \textbf{OCCURS clause}: allow KEY phrase and INDEXED phrase in any order.
\item \textbf{OCCURS clause}: fixed \bug{167}, where overly large numeric literals where accepted in the OCCURS clause.
\item \textbf{OCCURS clause (depending)}: require the minimum length to be less than the maximum length (\fr{99}).
\item \textbf{OCCURS clause (depending)}: disabled nested OCCURS DEPENDING tables due to bugs.
\item \textbf{OCCURS clause (screen-section)}: require relative LINE\slash{}COLUMN clauses in OCCURS entries (\bug{83}).
\item \textbf{OCCURS clause (unbounded)}: added by Frank Swarbrick (\patch{50}).
\item \textbf{PICTURE clause}: restricted number of permitted PICTURE strings (\bug{232}).
\item \textbf{PICTURE clause}: improved checks of constant-names referenced in PICTURE strings.
\item \textbf{RENAMES items}: strengthened syntax checks.
\item \textbf{RESERVE clause}: allow the optional word AREAS.
\item \textbf{Screen description}: permit figurative constants in screen items (\bug{108}).
\item \textbf{TALLY special register}: added.
\item \textbf{USAGE clause}: added ACUCOBOL's HANDLE phrases (see \fr{77}).
\item \textbf{VALUE clause}: VALUE clauses in REDEFINES entries now cause warnings, not errors, for compatibility.
\item \textbf{Variable records}: added checks that the minimum size of a variable record is large enough to contain the record key.
\end{itemize}

\paragraph{Procedure division}
\begin{itemize}
\item \textbf{ACCEPT statement}: added ESCAPE as synonym for EXCEPTION.
\item \textbf{ACCEPT statement}: permit clauses in any order.
\item \textbf{ACCEPT statement}: allow WITH before every screen attribute clause.
\item \textbf{ACCEPT statement}: entering control-C now terminates the program.
\item \textbf{ACCEPT statement (screen)}: fixed failed ACCEPTs caused by a buffer overflow.
\item \textbf{ACCEPT statement (screen)}: enhanced support for special keys (insert, tab, delete, alt-delete, etc.).
\item \textbf{ACCEPT statement (screen)}: fixed \bug{161} where screens terminated after entering a few characters in a field.
\item \textbf{ACCEPT statement (screen)}: added DEFAULT as synonym for UPDATE.
\item \textbf{ACCEPT statement (screen)}: ERASE and BLANK clauses in screens are now ignored (\bug{192}).
\item \textbf{ACCEPT statement (screen)}: fixed \bug{160} where ACCEPT statement LINE\slash{}COLUMN clauses did not work.
\item \textbf{ACCEPT statement (screen)}: fixed segfault on ACCEPT OMITTED (\bug{300}).
\item \textbf{ACCEPT statement (screen)}: added checks that screen attributes are not specified multiple times or after conflicting attributes.
\item \textbf{ACCEPT statement (screen)}: fixed some phrases not being recognised without being preceded by WITH (\bug{402}).
\item \textbf{ACCEPT statement (screen)}: fixed the backspace and delete keys not working and the insert key not toggling between insertion and overwriting.
\item \textbf{ACCEPT statement (screen)}: cursor now changes with insertion\slash{}overwrite mode (if supported by the terminal).
\item \textbf{ACCEPT statement (screen)}: a beep is emitted on attempts to \emph{insert} data into a full field.
\item \textbf{ADD statement (corresponding)}: restricted to numeric items (\bug{235}).
\item \textbf{ADD statement (table)}: added detection of ADD TABLE.
\item \textbf{Addition of COMP-3 numbers}: fixed bug where COMP-3 addition failed.
\item \textbf{Addition of floating-point numbers}: fixed incorrect addition of floating-point numbers.
\item \textbf{CALL statement}: implemented \fr{101}, allowing more arguments to be provided.
\item \textbf{CALL statement}: fixed behaviour when calling cancelled modules.
\item \textbf{CALL statement}: added RETURNING NOTHING. % TO-DO: What does this do?
\item \textbf{CALL statement}: the generation of C function declarations for static CALLs can now be disabled.
\item \textbf{CALL statement}: added checks for static CALLs referring to C macros.
\item \textbf{CALL statement}: warn if a literal containing the program-name contains spaces.
\item \textbf{CALL statement}: added detection of NESTED phrase.
\item \textbf{CANCEL statement}: fixed crash caused by cancelling a cancelled module.
\item \textbf{Conditions}: restricted use of IS (\bug{321}).
\item \textbf{Conditions}: added warnings for always true\slash{}false conditions (including the reason why it is always true\slash{}false).
\item \textbf{DESTROY statement}: added detection of DESTROY.
\item \textbf{DISPLAY statement}: permit clauses in any order.
\item \textbf{DISPLAY statement}: allow WITH before every screen attribute clause.
\item \textbf{DISPLAY statement (screen)}: fixed bug where EC-SCREEN exceptions did not trigger ON EXCEPTION handler (\bug{243}).
\item \textbf{DISPLAY statement (screen)}: fixed bugs in DISPLAY SPACES\slash{}ALL X``02''\slash{}ALL X``07''.
\item \textbf{DISPLAY statement (screen)}: added checks that screen attributes are not specified multiple times or after conflicting attributes.
\item \textbf{DISPLAY statement (screen)}: DISPLAY OMITTED marked as unfinished; currently equivalent to DISPLAY LOW-VALUE.
\item \textbf{DISPLAY statement (screen)}: fixed some phrases not being recognised without being preceded by WITH (\bug{402}).
\item \textbf{END DECLARATIVES phrase}: fixed \bug{88}, where an erroneous unreachable code warning was emitted for code without a main procedure.
\item \textbf{ENTRY statement}: suppress incorrect unreachable code warnings.
\item \textbf{Exception handlers}: permit NOT ON EXCEPTION\slash{}END-OF-PAGE\slash{}etc. before ON EXCEPTION\slash{}END-OF-PAGE\slash{}etc.
\item \textbf{EXIT statement}: added extension RETURNING\slash{}GIVING clause for PROGRAM phrase.
\item \textbf{File I-O}: added detection of and handling for error when no disc space is available for output files.
\item \textbf{File I-O}: added RETRY and ADVANCING ON LOCK as pending features.
\item \textbf{File I-O}: fixed detection of DISAM file handler.
\item \textbf{FREE statement}: NULL addresses no longer cause an exception.
\item \textbf{GOBACK statement}: added extension RETURNING\slash{}GIVING clause.
\item \textbf{INITIALIZE statement}: fixed \bug{84}, where literals could be passed to INITIALIZE.
\item \textbf{INITIALIZE statement}: fixed \bug{287}, where reference-modified group items were not treated like elementary items.
\item \textbf{INSPECT statement}: fixed \bug{47}, where clauses were permitted in invalid orders.
\item \textbf{LENGTH OF phrase}: fixed \bug{89}, where the length of REDEFINES item where calculated incorrectly.
\item \textbf{LENGTH OF phrase}: fixed \bug{110}, where LENGTH OF was not allowed in the UNTIL phrase of a PERFORM statement or in a VALUE clause.
\item \textbf{MOVE statement}: added more checks for overlapping MOVE statements.
\item \textbf{MOVE statement}: fixed truncation of COMP numbers not conforming to the binary-truncate setting (\bug{69}).
\item \textbf{MOVE statement}: fixed \bug{344}, where trying to MOVE to a procedure-name caused a segfault.
\item \textbf{MOVE statement}: added support for IBM's character-by-character MOVE.
\item \textbf{PERFORM statement}: fixed \bug{368}, where the compiler segfaulted when there was a PERFORM statement with an empty body and DEBUGGING MODE was specified.
\item \textbf{Procedure division header}: fixed \bug{55}, where a user-defined function without parameters failed to compile.
\item \textbf{Procedure division header}: disabled the BY VALUE phrase, pending a working implementation.
\item \textbf{Procedure division header}: fixed \bug{349}, where BY VALUE pointer parameters lead to code that couldn't be compiled by older versions of Microsoft Visual C++ (patched by Mario Matos).
\item \textbf{Procedure division header}: RETURNING items must now be declared in the linkage section.
\item \textbf{Procedure division header}: added RETURNING OMITTED.
\item \textbf{Procedure division header}: added entry-convention specifiers.
\item \textbf{Procedure division header}: now mandatory in function definitions (see \bug{271}).
\item \textbf{Procedure division header}: CHAINING programs may no longer be called by other programs (\bug{354}), per the ACUCOBOL implementation.
\item \textbf{Reference modification}: fixed \bug{146}, where the length of reference-modified item in an OCCURS DEPENDING table was too long because it was assumed the OCCURS DEPENDING table was at its maximum size.
\item \textbf{READ statement}: a failed second READ of a missing OPTIONAL file now results in a file status of 46, not 23.
\item \textbf{REWRITE statement}: added REWRITE FILE (\fr{170}).
\item \textbf{Screen I-O}: added detection of situations which raise EC-SCREEN-LINE-NUMBER, EC-SCREEN-STARTING-COLUMN and EC-SCREEN-ITEM-TRUNCATED.
\item \textbf{Screen I-O}: added support for the LINE 0 and COL 0 extensions.
\item \textbf{Screen I-O}: added some ACUCOBOL synonyms (NO ECHO, OFF, REVERSED, REVERSE, etc.).
\item \textbf{Screen I-O}: added detection of ACUCOBOL's non-standard clauses like TAB, NO-ECHO, STANDARD, BACKGROUND-HIGH, BACKGROUND-LOW, BACKGROUND-STANDARD and SIZE.
\item \textbf{SEARCH statement (ALL)}: fixed \bug{314}, where SEARCH ALL with an empty OCCURS DEPENDING table did not exit as soon as possible.
\item \textbf{Segment numbers}: added syntax checks.
\item \textbf{SET statement (address)}: disallowed changing address of non-01\slash{}77-level item (\bug{366}).
\item \textbf{SET statement (attribute)}: made HIGHLIGHT ON imply LOWLIGHT OFF and vice versa.
\item \textbf{SET statement (exception)}: added.
\item \textbf{SET statement (thread)}: added detection of ACUCOBOL extension.
\item \textbf{STOP statement (identifier)}: added (see \bug{320}).
\item \textbf{STOP statement (literal)}: fixed segfault.
\item \textbf{STOP statement (thread)}: added detection of ACUCOBOL extension.
\item \textbf{STRING statement}: strengthened syntax checks (\bug{259}).
\item \textbf{SUBTRACT statement (corresponding)}: restricted to numeric items (\bug{235}).
\item \textbf{SUBTRACT statement (table)}: added detection of SUBTRACT TABLE.
\item \textbf{Tracing}: fixed \bug{216}, where a segfault occurred with a program made from modules some of which had been compiled with tracing and physical CANCEL % TO-DO: Define "physical CANCEL"
enabled and some of which hadn't.
\item \textbf{UNSTRING statement}: fixed \bug{54}, where the POINTER value was calculated incorrectly when the delimiter was longer than one character.
\item \textbf{UNSTRING statement}: allow a literal to be the subject of an UNSTRING.
\item \textbf{WRITE statement}: added WRITE FILE (\fr{170}).
\end{itemize}

\paragraph{Intrinsic functions}
\begin{itemize}
\item \textbf{New functions (ACUCOBOL)}: ABSOLUTE-VALUE (synonym for ABS).
\item \textbf{New functions (COBOL 2014)}: FORMATTED-CURRENT-DATE, FORMATTED-DATE, FORMATTED-DATETIME, FORMATTED-TIME, INTEGER-OF-FORMATTED-DATE, TEST-FORMATTED-DATETIME.
\item \textbf{ISO-8601-date-handling functions}: added extension SYSTEM-OFFSET as replacement for last optional argument.
\item \textbf{ISO-8601-date-handling functions}: added EC-IMP-UTC-UNKNOWN if a time format ending in Z is provided but the timezone cannot be found.
\item \textbf{LENGTH function}: added detection of PHYSICAL phrase.
\item \textbf{RANDOM function}: fixed non-random number generation.
\end{itemize}

\paragraph{Built-in subprograms}
\begin{itemize}
\item \textbf{CBL\_GC\_FORK}: added.
\item \textbf{CBL\_GC\_PRINTABLE}: renamed from C\$PRINTABLE.
\item \textbf{CBL\_GC\_WAITPID}: added.
\item \textbf{CBL\_SET\_CSR\_POS}: added (feature requests \frnoname{148} and~\frnoname{187}).
\item \textbf{CBL\_READ\_KBD\_CHAR}: added (feature requests \frnoname{148} and~\frnoname{187}).
\end{itemize}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "grammar.tex"
%%% End:
