\chapter{Language element lists}

This appendix is required for conformance to the COBOL standard.

\section{Implementor-defined language element list}

\begin{enumerate}
\item ACCEPT statement (conversion of data).
\item ACCEPT statement (device used when FROM is unspecified).
\item ACCEPT statement, screen format (result when screen items overlap).
\item ACCEPT statement, screen format (when data is verified, behavior for inconsistent data).
\item ACCEPT statement (size of data transfer).
\item Alignment of alphanumeric group items (relative to first elementary item).
\item Alignment of data for increased efficiency (special or automatic alignment: interpretation, implicit filler,
  semantics of statements).
\item ALPHABET clause (ordinal number of characters in the native coded character sets).
\item Alphanumeric literals (number of hexadecimal digits that map to an alphanumeric character).
\item ASSIGN clause, USING phrase (meaning and rules for operands; consistency rules).
\item BACKGROUND-COLOR clause (the background color when the clause is not specified or the value specified is not in the range 0 to 7).
\item Byte (number of bits in).
\item CALL statement (rules for program-name formation for a non-COBOL program).
\item CALL statement (runtime resources that are checked).
\item CALL statement (rules for locating a non-COBOL program).
\item CALL statement (calling a non-COBOL program).
\item CALL statement (other effects of the CALL statement).
\item CANCEL statement (result of canceling an active program when EC-PROGRAM-CANCEL-ACTIVE is not
  enabled).
\item CANCEL statement (result of canceling a non-COBOL program).
\item Case mapping.
\item CHAR function (which one of the multiple characters is returned).
\item CHAR-NATIONAL function (which one of the multiple characters is returned).
\item Characters prohibited from use in text-words in COPY ... REPLACING and REPLACE statements.
\item CLOSE statement (closing operations).
\item COBOL character repertoire (encoding of, mapping of, substitute graphics).
\item COBOL character repertoire (if more than one encoding in a compilation group, control functions if any).
\item Color number (for a monochrome terminal, the mapping of the color attributes onto other attributes).
\item Compiler directives, compiler directive IMP (syntax rules and general rules).
\item Computer's coded character set (characters in and encoding of computer’s alphanumeric coded character set and computer’s national coded character set, encoding for usage DISPLAY and usage NATIONAL).
\item Computer's coded character set (coded character values for certain COBOL items).
\item Computer's coded character set (correspondence between alphanumeric and national characters).
\item Computer's coded character set (for literals, correspondence between compile-time and runtime character sets, when conversion takes place).
\item Computer's coded character set (correspondence between lowercase and uppercase letters when a locale is not in effect).
\item Computer's coded character set (when composite alphanumeric and national, mapping of characters to each).
\item Computer's coded character set (when more than one encoding, the mechanism for selecting encoding for runtime).
\item Computer's coded character set (whether UTF-8 or mixed alphanumeric and national characters recognized
  in class alphanumeric; applicable syntax and general rules).
\item COPY statement (rules for identifying and locating default library text).
\item CRT status 9xxx (the value of xxx for unsuccessful completion with implementor-defined conditions).
\item Cultural ordering table (allowable content of literal defines a cultural ordering table).
\item Currency symbol (equivalence of non-COBOL characters).
\item Currency symbol (implementor-defined prohibition of non-COBOL characters).
\item Cursor (the cursor movement if keys are defined that change the cursor position).
\item Data storage (possible representations when implementation provides multiple ways of storing data).
\item Default encoding specifications (for standard decimal floating-point usages).
\item Default endianness specifications (for standard floating-point usages).
\item DEFINE directive (mechanism for providing value of a compilation-variable-name from the operating environment).
\item Devices that allow concurrent access.
\item DISPLAY statement (data conversion).
\item DISPLAY statement (format for display of a variable-length group).
\item DISPLAY statement (size of data transfer).
\item DISPLAY statement (standard display device).
\item Dynamic-capacity table (determination of highest permissible occurrence number).
\item Dynamic-capacity table (physical allocation).
\item Dynamic-length elementary items (maximum length).
\item Dynamic-length elementary items (structure if dynamic-length-structure-name-1 is not specified).
\item ENTRY-CONVENTION clause (entry-convention-names, their meanings and the default when not specified).
\item EXIT and GOBACK statements (execution continuation in a non-COBOL runtime element).
\item Exponentiation (results for certain operand values).
\item External repository (mechanism for specifying whether checking and updating occur).
\item External repository information (other information beyond the required information).
\item Externalized names (formation and mapping rules).
\item Fatal exception condition (whether detected at compile time, circumstances under which detected).
\item Fatal exception condition (whether or not execution will continue, how it will continue, and how any receiving operands are affected when events that would cause a fatal exception to exist occur but checking for that condition is not enabled).
\item FILE-CONTROL entry, ASSIGN clause (TO phrase meaning and rules).
\item FILE-CONTROL entry, ASSIGN clause (consistency rules for external file connectors).
\item FILE-CONTROL entry, ASSIGN clause (USING phrase meaning and rules).
\item Figurative constant values (representation of zero, space, and quote).
\item File sharing (interaction with other facilities and languages).
\item File sharing (which devices allow concurrent access to the file).
\item File sharing (default mode when unspecified).
\item Fixed file attribute (whether the ability to share a file is a fixed file attribute).
\item FLAG-02 directive (warning mechanism).
\item FLAG-85 directive (warning mechanism).
\item Floating-point numeric item (alignment when used as a receiving operand).
\item Floating-point numeric literals (maximum permitted value and minimum permitted value of the exponent).
\item FOREGROUND-COLOR clause (the foreground color when the clause is not specified or the value specified is
  not in the range 0 to 7).
\item FORMAT clause (representation produced).
\item FORMAT clause (exclusions on restoring to same internal representation).
\item Format validation (rules for checking items of usages other than display or national).
\item FORMATTED-CURRENT-DATE (accuracy of returned time).
\item Function-identifier (execution of a non-COBOL function when a function-prototype-name is specified).
\item Function-identifier (object time resources that are checked).
\item Function-identifier (result when argument rules are violated and checking for the EC-ARGUMENT-FUNCTION
  exception condition is not enabled).
\item Function keys (context-dependent keys, function number, and method for enabling and disabling).
\item Function returned values (characteristics, representation, and returned value for native arithmetic).
\item Hexadecimal alphanumeric literals (mapping for non-existing corresponding character).
\item Hexadecimal alphanumeric literals (mapping when characters not multiples of four bits).
\item Hexadecimal national literals (mapping for non-existing corresponding character).
\item Hexadecimal national literals (mapping when characters not multiples of four bits).
\item Implementor-defined exception conditions, EC-IMP-xxx (specification and meaning of xxx).
\item Implementor-defined level-2 exception conditions, EC-level-2-IMP (specification and meaning of the specified level-2 exception condition).
\item INVOKE statement (behavior when invoking a non-COBOL method).
\item INVOKE statement (runtime resources that are checked).
\item I-O status (action taken for fatal exception conditions).
\item I-O status (if more than one value applies).
\item I-O status, permanent error (technique for error correction).
\item I-O status 0x (value of x).
\item I-O status 24 (manner in which the boundaries of a file are defined).
\item I-O status 34 (manner in which the boundaries of a file are defined).
\item I-O status 52 (conditions under which deadlock is detected).
\item I-O status 9x (value of x).
\item LEAP-SECOND directive (whether a value greater than 59 seconds may be reported and, if so, the maximum
  number of seconds that may be reported).
\item LEAP-SECOND directive (whether standard numeric time form values greater than or equal to 86,400 may be
  reported).
\item Life cycle for objects (timing and algorithm for taking part in continued execution).
\item Linkage section (whether access to linkage section items is meaningful when called from a non-COBOL
  program).
\item Listings (whether and when produced by the compiler, effect of logical conversion).
\item Locale specification (how user and system defaults defined; at least one user and one system default).
\item Locale specification (manner of implementation).
\item Locale switch (whether a switch by a non-COBOL runtime module is recognized by COBOL).
\item METHOD-ID paragraph (actual method-name used when PROPERTY phrase is specified).
\item National literals (number of hexadecimal digits that map to a national character).
\item Native arithmetic (techniques used, intermediate data item).
\item Native arithmetic (when an operand or arithmetic expression is an integer).
\item NULL (value of NULL).
\item OBJECT-COMPUTER paragraph (default object computer).
\item OBJECT-COMPUTER paragraph (computer-name and implied equipment configuration).
\item OCCURS clause (range of values allowed in the index).
\item OPEN statement (validation of fixed file attributes).
\item OPEN statement with OUTPUT phrase (positioning of the output file with regard to physical page boundaries).
\item OPEN statement without the SHARING phrase and no SHARING clause in the file control entry (definition of sharing mode established for each file connector).
\item Parameterized classes and interfaces (when expanded).
\item Procedure division header rules when either the activating or the activated runtime element is not a COBOL element (restrictions and mechanisms for all supported language products with details such as the matching of parameters, data type representation, returning of a value, and omission of parameters).
\item Program-address identifier (relation between address and non-COBOL program).
\item Program-name (formation rules for a non-COBOL program).
\item RANDOM function (seed value when no argument on first reference).
\item RANDOM function (subset of the domain of argument-1).
\item RECORD clause (calculations to derive size of records on storage medium).
\item RECORD clause (implicit RECORD clause if RECORD clause is not specified).
\item RECORD clause (whether fixed or variable records produced for fixed-or-variable-length format).
\item RECORD DELIMITER clause (consistency rules when used with external file connectors).
\item RECORD DELIMITER clause (feature-name and associated method for determining length of variable-length
  records).
\item RECORD DELIMITER clause (if not specified, method for determining length of variable-length records).
\item Record locking (circumstances other than a locked logical record that return a locked record status).
\item Record locking (default mode when unspecified by user).
\item Record locking (maximum number allowed for a run unit).
\item Record locks (maximum number allowed for a file connector).
\item Reference format (control characters in a free-form line).
\item Reference format (meaning of lines and character positions in free-form and fixed-form format).
\item Reference format (rightmost character position of program-text area).
\item Report file (record structure).
\item Report writer printable item (fixed correspondence between columns and national characters).
\item REPOSITORY paragraph (how external repository and class-specifier determine which class is used).
\item REPOSITORY paragraph, INTERFACE phrase (how interface specifier and external repository determine
  which interface is used).
\item REPOSITORY paragraph (when the AS phrase is required).
\item RESERVE clause (number of input-output areas, if not specified).
\item RETRY phrase (interval between attempts to obtain access to a locked file or record).
\item RETRY phrase (maximum meaningful time-out value and internal representation; technique for determining
  frequency of retries).
\item Run unit (relationship and interaction with non-COBOL components).
\item Run unit termination (whether locale reset).
\item SAME SORT/SORT-MERGE AREA clause (extent of allocation).
\item SEARCH ALL statement (varying of the search index during the search operation).
\item SECONDS-PAST-MIDNIGHT function returned value (precision).
\item SECURE clause (cursor movement when data is entered into a field for which the SECURE clause is specified).
\item SELECT WHEN clause (whether a SELECT WHEN takes effect for READ statements and REWRITE or WRITE
  statements with the FILE phrase in the absence of a CODE-SET clause or a FORMAT clause).
\item SET statement (effect of SET on function whose address is being stored in a function-pointer).
\item SET statement (effect of SET on program whose address is being stored in a program-pointer).
\item SET statement (value in NaN payload).
\item SIGN clause (representation when PICTURE contains character 'S' with no optional SIGN clause).
\item SIGN clause (valid sign when SEPARATE CHARACTER phrase not present).
\item Size error condition (whether or not range of values allowed for the intermediate data item is to be checked).
\item SPECIAL-NAMES paragraph (allowable locale-names and literal values).
\item SPECIAL-NAMES paragraph, ALPHABET clause (coded character set referenced by STANDARD-2 phrase).
\item SPECIAL-NAMES paragraph, ALPHABET clause, code-name-1 (alphanumeric coded character set and
  collating sequence; ordinal number of characters; correspondence with native alphanumeric character set).
\item SPECIAL-NAMES paragraph, ALPHABET clause, code-name-1 and code-name-2 (the names supported for
  code-name-1 and code-name-2).
\item SPECIAL-NAMES paragraph, ALPHABET clause, code-name-2 (national coded character set and collating
  sequence; ordinal number of characters; correspondence with native national character set).
\item SPECIAL-NAMES paragraph, ALPHABET clause, literal phrase (ordinal number of characters not specified).
\item SPECIAL-NAMES paragraph, ALPHABET clause, STANDARD-1 and STANDARD-2 phrases (correspondence
  with native character set).
\item SPECIAL-NAMES paragraph, ALPHABET clause, UCS-4, UTF-8, and UTF-16 phrases (correspondence with
  native character set).
\item SPECIAL-NAMES paragraph, device-name (names available, restrictions on use).
\item SPECIAL-NAMES paragraph, feature-name (names available, any positioning rules, any restrictions on use).
\item SPECIAL-NAMES paragraph, switch-name (names available, which switches may be referenced by the SET
  statement, scope of, and external facility for modification).
\item Standard intermediate data item (representation).
\item STOP statement (constraints on the value of the STATUS literal or on the contents of the data item referenced
  by the STATUS identifier).
\item STOP statement (mechanism for error termination).
\item Subscripts (mapping indexes to occurrence numbers).
\item Switch-name (identifies an external switch).
\item SYNCHRONIZED clause (effect on elementary items and containing records or groups; implicit filler
  generation).
\item SYNCHRONIZED clause (how records of a file are handled).
\item SYNCHRONIZED clause (positioning when neither RIGHT or LEFT is specified).
\item System-names (rules for formation of a system-name).
\item Terminal screen (correspondence of a column and a character in the computer’s national coded character set).
\item Text manipulation (stage of processing the LISTING and PAGE directives and the SUPPRESS phrase of COPY).
\item Text manipulation (stage of processing parameterized class expansion).
\item Time formats and corresponding function values (maximum precision not less than nine fractional digits).
\item THROUGH phrase in VALUE clause and EVALUATE statement (collating sequence used for determining range
  of values when no alphabet-name is specified).
\item TURN directive (whether location information is available when the LOCATION phrase is not specified).
\item USAGE BINARY clause (computer storage allocation, alignment and representation of data).
\item USAGE BINARY-CHAR, BINARY-SHORT, BINARY-LONG, BINARY-DOUBLE (allow wider range than minimum
  specified).
\item USAGE BINARY-SHORT, BINARY-LONG, BINARY-DOUBLE, FLOAT-SHORT, FLOAT-LONG, FLOAT-EXTENDED (representation and length of data item associated with).
\item USAGE COMPUTATIONAL clause (alignment and representation of data).
\item USAGE DISPLAY (size and representation of characters).
\item USAGE FLOAT-SHORT, FLOAT-LONG, FLOAT-EXTENDED (size and permitted range of value).
\item USAGE FUNCTION-POINTER clause (alignment, size, and representation of data; and allowable languages).
\item USAGE INDEX clause (alignment and representation of data).
\item USAGE NATIONAL (size and representation of characters).
\item USAGE OBJECT REFERENCE clause (amount of storage allocated).
\item USAGE PACKED-DECIMAL clause (computer storage allocation, alignment and representation of data).
\item USAGE POINTER clause (alignment, size, representation, and range of values).
\item USAGE PROGRAM-POINTER clause (alignment, size, and representation of data; and allowable languages).
\item USE statement (action taken following execution of the USE procedure when I-O status value indicates a fatal exception condition).
\item User-defined words (whether extended letters may be specified in user-defined words externalized to the operating environment).
\item Variable-length data items (actual time when the resources used are freed).
\item WRITE statement (mnemonic-name-1).
\item WRITE statement (page advance when mnemonic-name-1 specified).
\end{enumerate}

\section{Optional language element list}

\subsection{ACCEPT and DISPLAY screen handling}

Complete support is claimed.

\subsection{ARITHMETIC IS STANDARD}

No support is claimed.

\subsection{Dynamic capacity tables}

The syntax is recognised, but no functionality is claimed.

\subsection{DYNAMIC LENGTH elementary items}

No support is claimed.

\subsection{Extended letters}

No support is claimed.

\subsection{File sharing and record locking}

Support is claimed, but the level of support is processor-dependent.

\subsection{FORMAT and SELECT WHEN file handling}

No support is claimed.

\subsection{Locale support and related functions}

The syntax is recognised, but no functionality is claimed.

\subsection{Object orientation}

No support is claimed.

\subsection{Report writer}

The syntax is recognised, but no functionality is claimed.

\subsection{RESUME statement}

No support is claimed.

\subsection{REWRITE FILE and WRITE FILE}

Complete support is claimed.

\subsection{VALIDATE}

No support is claimed.

\section{Non-standard extension list}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "grammar.tex"
%%% End:
