@section Intrinsic Python scripting

@*
Python scripting is an optional build feature of GnuCOBOL.
@*
Requires Python and @code{./configure --with-python[=PYTHON]} during build.
The @code{=PYTHON} part specifies a Python interpreter to run to determine
version information.

@*
Adds a new Intrinsic function, @code{FUNCTION PYTHON(script [,arg,...])}.
@*

This function accept character data as the script, unlimited arguments
(also character data), and returns an ALPHANUMERIC field from the Python
interpreter.  Reference modification of the result is allowed.

The @code{result} variable from within Python determines the value returned to
GnuCOBOL.

The Python dictionary is persistent across invocations of the intrinsic
function. To reset Python, use @code{FUNCTION PYTHON(0)} which is equivalent
to @code{FUNCTION PYTHON(PYAPI-FINALIZE)} if the pyapi.cpy copybook is
included.

Includes a @code{pyapi.cpy} copybook that defines special control codes, error
number constants and an @code{EXTERNAL} variable, @code{SCRIPT-RETURN-CODE}
that provides status values.

@code{pyapi.cpy} will be installed to a system specific location that is part
of the default search path for the compiler @code{COPY} directive.

@section Python examples

@verbatim
    move python("import sys; result = int(sys.argv[1] * int(sys.argv[2]", 6, 7)
      to answer
@end verbatim

Parameter 1 is the script, all other arguments (unlimited) are passed to Python
as @code{sys.argv} values. @code{sys.argv[0]} is "GnuCOBOL". Return value is
ALPHANUMERIC.

Reference modification is allowed.

    @code{display python("result = 'abc' * 5")(1:10)}

As this is defined as an alphanumeric function, use @code{FUNCTION NUMVAL} to
use any Python results in COBOL computations.

    @code{compute answer = numval(python("result = 6 * 6)) + 6}

@section Optional build feature

This is an optional build feature.

@code{./configure --with-python=PYTHON} to build a compiler with Python
scripting.  Supports Python 2.6 and up, including Python 3.

Requires a Python shared library and headers that match the configured build.
